---
title: "European Covid Comparisons Part 2"
output: html_notebook
---

This notebook continues the analysis I started in `european_covid_comparisons_1.Rmd`. In part 1 I wrangled the data and created a dataset of European countries' populations and gdp per capita. Part 2 will supplement this dataset with data on Covid-19 testing and deaths.

Data on Covid deaths is taken from [World Health Organization Situation Reports](https://www.who.int/docs/default-source/coronaviruse/situation-reports/20200328-sitrep-68-covid-19.pdf?sfvrsn=384bc74c_2)


## Import and inspect the dataset
```{r}
library(dplyr)
library(readr)
library(stringr)

pop_and_gdp <-read_csv('european_pop_gdp_per_capita.csv')
pop_and_gdp
```

## import covid-19 data from WHO stats
Source: [Situation Report 68](https://www.who.int/docs/default-source/coronaviruse/situation-reports/20200328-sitrep-68-covid-19.pdf?sfvrsn=384bc74c_2), which collated data from national authorities submitted by 10:00 CET 28 March 2020.
```{r}
covid_deaths <- read_csv('european_covid_deaths_2020_28_mar.csv')
covid_deaths
```

## Import data on Covid-19 testing
This is going to be complicated. WHO does not produce statistics on this.
```{r}
covid_tests <- read_csv('european_covid_tests_2020_29_mar.csv')
covid_tests
```


## combine the datasets
```{r}
master <- pop_and_gdp %>% bind_cols(covid_deaths) %>% bind_cols(covid_tests)
master <- master %>% select(-Country1,-X11,-Country2,-X12)
master
```

## Sort the data 
```{r}
sort_by_deaths <- master %>% arrange(desc(deaths))
sort_by_deaths
```

## Visualise the data

### a. Is there a correlation between GPD and testing?
    
```{r}
library(ggplot2)
library(scales)

change_ireland <- master %>% mutate(tests=replace(as.numeric(tests), X1==c(15), c(17992)))
change_spain <- change_ireland %>% mutate(tests=replace(as.numeric(tests), X1==c(27), c(355000)))
change_sweden <- change_spain %>% mutate(tests=replace(as.numeric(tests), X1==c(28), c(24500)))
numerics <- change_sweden %>% filter(tests!="not known")
numerics
#gdp_to_tests <- ggplot(master, aes(x=gdp_per_capita, y=tests)) + geom_line() 
#gdp_to_tests

gdp_to_tests <- ggplot(numerics,aes(x=gdp_per_capita)) + geom_point(aes(y=tests)) +  scale_y_continuous(labels=label_comma()) + labs(title="Plotting GDP per Capita and Number of Tests", x="GDP per Capita", y="Number of Tests")

 gdp_to_tests 


```

###  b. Is there a correlation between testing and deaths?
```{r}
tests_deaths <- ggplot(numerics,aes(x=tests)) + geom_point(aes(y=deaths)) + scale_x_continuous(labels=label_comma()) + labs(title="Plotting Number of Tests and Number of Deaths", x="Number of Tests", y="Number of Deaths")

tests_deaths
```


###  c. Is there a correlation between GDP and deaths?
```{r}
gdp_to_deaths <- ggplot(numerics,aes(x=gdp_per_capita)) + geom_point(aes(y=deaths)) +  scale_y_continuous(labels=label_comma()) + labs(title="Plotting GDP per Capita and Number of Deaths", x="GDP per Capita", y="Number of Deaths")

 gdp_to_deaths

```

## Using rates for measuring deaths and tests
Given the differences in population size, it could be misleading to compare the raw numbers of tests and deaths. It is expected that- other things being equal - larger populations will require more testing than smaller populations. So, I will calculate the per capita testing and death rates. Note these are linked to each country's total population, not the number of infected people.

```{r}
# add percent_tested
# This is the percentage of the population that has been tested.
#master <- master %>% mutate(tests=as.numeric(tests))

results_by_tests <- numerics %>% mutate(percent_tested=(tests/population)*100)
results_by_tests <- results_by_tests %>% mutate(percent_tested=format(percent_tested, digits=1))
results_by_tests <- results_by_tests %>% arrange(desc(percent_tested))
results_by_tests <- results_by_tests %>% mutate(percent_tested=as.numeric(percent_tested))
results_by_tests

```

## Is there a correlation between gpd_per_capita and tests_per_capita?
```{r}
gdp_to_tests_per_capita <- ggplot(results_by_tests,aes(x=gdp_per_capita)) + geom_point(aes(y=percent_tested)) +  scale_y_continuous(labels=label_comma()) + labs(title="Plotting GDP per Capita and Percentage of People Tested", x="GDP per Capita", y="Percentage of population tested")
gdp_to_tests_per_capita
```


```{r}
# add risk of death (%)
results_by_deaths <- results_by_tests %>% mutate(death_risk=(deaths/population)*100)
results_by_deaths <- results_by_deaths %>% mutate(death_risk=format(death_risk, digits=2, scientific=FALSE))
results_by_deaths <- results_by_deaths %>% arrange(desc(death_risk))
results_by_deaths <- results_by_deaths %>% mutate(death_risk=as.numeric(death_risk))
results_by_deaths

# I learned how to format a number with scientific=FALSE - this trnasformed a number 9.0e-04 into something I can understand (0.015136)
```

```{r}
gdp_to_death_risk <- ggplot(results_by_deaths,aes(x=gdp_per_capita)) + geom_point(aes(y=death_risk)) +  scale_y_continuous(labels=label_comma()) + labs(title="Plotting GDP per Capita and the Risk of Death", x="GDP per Capita", y="Risk of death (%)")
gdp_to_death_risk
```



```{r}
# Work out the natural frequency in a thousand. E.g. 3 people in every thousand will get a test. 
# define the function
frequency_in_a_thousand <- function(risk) {(risk/100)*1000}
# plug in some numbers

#all_frequencies <- function() { 
#for (row in 1:nrow(results_by_deaths)) {
#freq <- frequency_in_a_thousand(results_by_deaths[row, "death_risk"])
#country <- results_by_deaths[row, "Country"]
#speech <- str_glue("In {country}, {freq} people in every thousand have died of the virus.")
#print(speech)}
#  return()
#}

#bob <- all_frequencies()
#bob
# so, this function works, but it adds an unwanted NULL line. I'll try it without the wrapper function.
 
for (row in 1:nrow(results_by_deaths)) {
freq <- frequency_in_a_thousand(results_by_deaths[row, "death_risk"])
country <- results_by_deaths[row, "Country"]
speech <- str_glue("In {country}, {freq} people in every thousand have died of the virus.")
print(speech)}
 
#Note a problem with the frequency of Czech Republic, Croatia, Poland and Bulgaria. The format is "8e-04". The cause was the format of death_risk. If digits=1, the format is incorrect, but if I change it to 2, the format is correct.

```

## Express the frequency as a "1 in XXXX" chance
```{r}
frequency <- function(population, deaths) {
  return(population/deaths)
}

#for (row in 1:nrow(results_by_deaths)) {
#freq <- frequency(results_by_deaths[row, "population"], results_by_deaths[row, "deaths"])
#freq <- format(freq, big.mark = ",", digits=0, scientific = FALSE)
#country <- results_by_deaths[row, "Country"]
#speech <- str_glue("In {country}, 1 person in  {freq} has died of the virus.")
#print(speech)}

# Note that the for loop doesn't work for countries with zero deaths. It says that 1 person in Inf has died. Solution is to add an if statement to exclude 
for (row in 1:nrow(results_by_deaths)) {
  country <- results_by_deaths[row, "Country"]
  if (results_by_deaths[row, "deaths"] > 0) {
freq <- frequency(results_by_deaths[row, "population"], results_by_deaths[row, "deaths"])
freq <- format(freq, big.mark = ",", digits=0, scientific = FALSE)

speech <- str_glue("In {country}, 1 person in  {freq} has died of the virus.")
print(speech)}
  else {
    nil_speech <- str_glue("No one has died in {country} yet, so no frequency can be determined.")
    print(nil_speech)
  }
}

# I learned how to format the numbers with commas, using big.mark

```


---
# Sources
Where possible, data have been taken from official government websites. Where this was not possible, I have used a [Wikipedia page on Covid-19 testing](https://en.wikipedia.org/wiki/COVID-19_testing). 


[Austria](https://www.sozialministerium.at/Informationen-zum-Coronavirus/Neuartiges-Coronavirus-(2019-nCov).html)

[Bulgaria: Wikipedia page](https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_Bulgaria)
[Cyprus](https://www.pio.gov.cy/coronavirus/en/index.html)
[Hungary](https://koronavirus.gov.hu/)
[Iceland](https://www.covid.is/tolulegar-upplysingar)
[Lithuania](http://sam.lrv.lt/lt/naujienos/koronavirusas)
[Norway](https://www.fhi.no/en/id/infectious-diseases/coronavirus/daily-reports/daily-reports-COVID19/)

[Romania](https://www.mai.gov.ro/informare-covid-19-grupul-de-comunicare-strategica-29-martie-ora-13-00/)

[Slovenia](https://www.korona.gov.sk/)

[Sweden](https://www.folkhalsomyndigheten.se/smittskydd-beredskap/utbrott/aktuella-utbrott/covid-19/aktuellt-epidemiologiskt-lage/)



