---
title: "Comparison of UK and Sweden"
output: html_notebook
---

The reason I am comparing the UK and Sweden is that both countries initially took a minimalist approach to tackling the coronavirus. That is, they stopped short of lockdown. However, on 23rd March the UK switched strategy and went into lockdown. Sweden, by contrast, continued with their minimalist approach. So, I am interested to find out whether these different approaches resulted in a statistically significant difference in the number of deaths caused by Covid-19.

## Hypothesis
The UK death toll per 100,000 people is lower than in Sweden.

## Null hypothesis
uk_death_rate >- sweden_death_rate
 

```{r}
library(stringr)
library(dplyr)
library(readr)
```

## Import the swedish dataset and add a column for deaths as a percentage of the population
```{r}
swedish_population <- 10322000

sweden <- read.csv("covid_19_sweden_stats.csv")
sweden <- sweden %>% mutate(sweden_death_rate=format(100*(Deaths/swedish_population), scientific=FALSE))

sweden
```

## Import the UK dataset and add a column for deaths as a percentage of the population
```{r}
uk_population <- 66867000
uk <- read_csv("covid_19_deaths.csv")
uk <- uk %>% select(-`Global Deaths`)
uk <- uk %>% rename(deaths='UK Deaths')
uk <- uk %>% mutate(uk_death_rate=format(100*(deaths/uk_population), scientific=FALSE))
uk
```

## Merge the datasets and compare
```{r}
uk_sweden <- sweden %>% bind_cols(uk)
uk_sweden <- uk_sweden %>% select(Date, sweden_death_rate, uk_death_rate)
uk_sweden <- uk_sweden %>% mutate(sweden_death_rate=as.numeric(sweden_death_rate), uk_death_rate=as.numeric(uk_death_rate))

uk_sweden <- uk_sweden %>% mutate(comparison=uk_death_rate/sweden_death_rate)
uk_sweden <- uk_sweden %>% mutate(sweden_death_rate=format(sweden_death_rate, scientific=FALSE), uk_death_rate=format(uk_death_rate, scientific=FALSE))
uk_sweden


```
The comparison column shows how much bigger the uk_death_rate is than the sweden_death_rate. So, if the comparison column is 1.534228, then the uk_death_rate is just over 53% higher than the sweden_death_rate.

```{r}
#calculate the average death rates (mean and median - the values are too specific for the mode to be useful

#mean
swedish_mean_death_rate <- uk_sweden %>% summarize(mean(as.numeric(sweden_death_rate)))
str_glue("Swedish mean death rate: {swedish_mean_death_rate}")

uk_mean_death_rate <- uk_sweden %>% summarize(mean(as.numeric(uk_death_rate)))
str_glue("UK mean death rate: {uk_mean_death_rate}")

mean_comparison <- uk_mean_death_rate/swedish_mean_death_rate
str_glue("When comparing mean death rates, the UK death rate is {mean_comparison} times the Swedish death rate.")

#median
swedish_median_death_rate <- uk_sweden %>% summarize(median(as.numeric(sweden_death_rate)))
uk_median_death_rate <- uk_sweden %>% summarize(median(as.numeric(uk_death_rate)))
median_comparison <- uk_median_death_rate/swedish_median_death_rate
swedish_median_death_rate <- format(swedish_median_death_rate, scientific=FALSE)
uk_median_death_rate <- format(uk_median_death_rate, scientific=FALSE)
str_glue("Swedish median death rate: {swedish_median_death_rate}")


str_glue("UK median death rate: {uk_median_death_rate}")


str_glue("When comparing median death rates, the UK death rate is {median_comparison} times the Swedish death rate.")






```


## Plot the comparison 
```{r}
library(ggplot2)
library(tidyr)
install.packages("tidyverse")
library("tidyverse")
library(scales)
df <- uk_sweden %>% gather(key = "variable", value = "value", -Date)
#head(df)
ggplot(df, aes(x = Date, y = value)) + 
  geom_line(aes(color = variable, linetype = variable)) + 
  scale_color_manual(values = c("darkred", "steelblue"))

#comparison_line <- ggplot(df, aes(x=Date)) + geom_line(aes(y=sweden_death_rate), color="blue") + geom_line(aes(y=uk_death_rate), color='red') + geom_text(aes(x=as.Date("2020-03-01"), y=0.00000000000, label="Sweden")) + geom_text(aes(x=as.Date("2020-03-25"), y=0.0022, label="UK"))
#comparison_line

#psavert_label <- "PSAVERT"
#uempmed_label <- "UEMPMED"
#ggplot(economics, aes(x=date)) + 
#  geom_line(aes(y = psavert), color = "darkred") +
#geom_line(aes(y = uempmed), color="steelblue", linetype="twodash") + geom_text(aes(x=as.Date("1968-10-01"), y = 15, label=psavert_label)) + #geom_text(aes(x=as.Date("2013-10-01"), y = 25, label=uempmed_label)) + labs(title="US Personal Savings Rate and Median Weeks Unemployed, #1967-2015")


```



