---
title: "Using the Predictor: Daily Growth Rate of 1.25"
output: html_notebook
---

This code uses the file `predictions.csv` created in `create_predictions.Rmd`. 

## My hypothesis:
Deaths are increasing at a daily rate of 1.25. That is, each day, the cumulative number of deaths increases by a quarter. 

## How will I test this?
* Compare the value in predicted_deaths with actual_deaths. The figure of actual_deaths is taken from the World Health Organization's [daily situation reports](https://www.who.int/emergencies/diseases/novel-coronavirus-2019/situation-reports/) on the number of deaths. 

## What would constitute failure of the hypothesis?
I'm not expecting the daily rate to be exactly 1.25, but if the predicted_deaths values are more than 10% greater/lower than actual_deaths values, that would indicate the hypothesis is incorrect. However, it is expected that more data will allow me to amend my hypothesis.

```{r}
library(stringr)
library(readr)
library(dplyr)
```


```{r}
predictions <- read_csv('predictions.csv')
predictions
```

```{r}
# change actual_deaths to numeric
predictions <- predictions %>% mutate(actual_deaths=as.numeric(actual_deaths))
predictions
```


```{r}
# update the actual_deaths of X1
predictions <- predictions %>% mutate(actual_deaths=replace(as.numeric(actual_deaths), X1==1, 578))
predictions
```

```{r}
# add a measure of the distance between actual_deaths and predicted_deaths
predictions_with_distance <- predictions %>% mutate(distance_as_percent=((predicted_deaths-actual_deaths)/predicted_deaths)*100)
predictions_with_distance <- predictions_with_distance %>% mutate(distance_as_percent=format(distance_as_percent, digits=2))
predictions_with_distance
```

```{r}
#calculate true growth rate - I'm not sure that my formula is correct
predictions_with_true_growth_rate <- predictions_with_distance %>% mutate(true_growth_rate=(1+(actual_deaths-463)/463))
predictions_with_true_growth_rate <- predictions_with_true_growth_rate %>% mutate(true_growth_rate=format(true_growth_rate, digits=2))
predictions_with_true_growth_rate
```




